\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-08-16 gwnnDraft02.tex
Lots of draftnotes.

\title{\LARGE{{\bf{Piled up gravitational waves:} \\
\bf{Searching for new signals of $N$naturalness} }}}
\author{{\bf {Paul Archer-Smith, Dylan Linthorne, and Daniel Stolarski}}}

@@

\section{Gravitational Waves}
\label{sec:gw}
Sectors other than the SM can be reheated through multifield inflationary theories, Reheaton models, etc. In these models, the SM will receive the majority of the energy density while the remaining energy density will be distributed, in a model dependent way, to the various hidden sectors. If there does not exist a portal between the sectors, they will remain decoupled and evolve independently. When looking at multiple thermally decoupled hidden sectors it is convienent to define the temperature ratio [cite cold,dark,noisy]
\begin{equation}
\xi_i \equiv  \frac{T_{i}}{T_{\gamma}}
\end{equation}
Where $T_{i}$ is the temperature of the $i^{th}$ hidden sector, and $T_{\gamma}$ the temperature of the standard model (assuming radiation domination).

The amount of energy density dumped into the individual sectors dictates the energy budget for the PT and hence for the gravitational waves.  A quantity that characterizes the strength of the PT is the ratio of the latent heat and energy density of radiation, at the time of nucleation, 
\begin{equation}
\alpha = \frac{\epsilon}{g_{*} \pi^2 (T^{nuc}_{\gamma})^4/30}
\end{equation}
Assuming that there is a negligible amount of energy being dumped back into the SM radiation bath, which would cause significant reheating of $\rho_{\gamma}$, the latent heat of the PT $\epsilon$ should  approximately be the energy density of the hidden sector going through the PT. $g_{*}$ has weak temperature dependence in a single sector, but when dealing with multiple hidden sectors, $g_{*}$ gains contributions from all $N$ sector's relativistic degrees of freedom, weighted by their respective temperature ratios (energy densities).
\begin{equation}
g_{*} = g_{*,\gamma} + \sum_{i = 1}^{N} g_{*,i} (\xi_{i})^4
\end{equation}
It will be seen that for models with a significant number of hidden sectors, that the lack of substantial reheating in adjecent sectors will cause their respective $g_{*}$ corrections to become negligible, $g_{*} \approx g_{*,\gamma}$.
\section{Particle Setup}
\label{sec:nn}

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=0.4\textwidth]{standardDecay.eps}
\includegraphics[width=0.4\textwidth]{exoticDecay.eps}
\end{minipage}
\hfill
\caption{Feynman diagrams for primary reheaton decays assuming $m_{\phi} \ll m_H$. The left decay is the primary decay for standard sectors ($\langle H\rangle\neq 0$), with $f$ representing all possible fermions. The right decay occurs in exotic sectors ($\langle H\rangle = 0$)
}
\label{fig:reheatonDecays}
\end{figure*}

@@

Keeping with the original convention, we write our mass parameter as
\begin{equation}\label{eqn:massParam}
\left(m_H^2\right)_i = -\frac{\Lambda_H^2}{N}(2i+r),
\end{equation}
where $i$ denotes the sector ($-\frac{N}{2}\leq i \leq \frac{N}{2}$). $i = 0$ is the sector that possess the smallest vev and is thus ``our" SM sector. Finally, the parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:signals}.

@@

A key issue within $N$naturalness is how to predominantly gift energy density to our own sector so as to not be immediately excluded by number of effective neutrinos ($N_{eff}$) bounds. This is done through the introduction of a post-inflationary field called the ``reheaton". After inflation, the reheaton field possess the majority of the energy density of the Universe. Although this field can generically be either by bosonic or fermionic, we reduce our scope to a scalar reheaton $\phi$. Our focus is primarily the production of gravitational waves from multiple sectors: the shifting to a fermion reheaton doesn't change the scaling of the energy density of the exotic sectors and, as a result, is ignored in our analysis.

@@

Ultimately, after examining the gravitational wave case produced by standard $N$naturalness, we generalize our results to include the (rather unnatural) case where sectors are reheated ``randomly" as opposed to in relation to the higg's mass parameter. This allows us to explore more general models with multiple dark sectors at a huge range of scales. For these models, the reheating mechanism remains undefined and the results of this section are completely irrelevant.

The appropriate Lagrangian for a scalar reheaton $\phi$ is: 
\begin{equation}\label{eqn:nnLagrangian}
\cL_\phi \supset -a\phi\sum_i\vert H_i\vert^2 - \frac{1}{2}m^2_\phi \phi^2.
\end{equation}
Note that the cross-quartic couplings are absent, suppressed by a very small coupling. Effective Lagrangians for the two different types of sectors present in this theory can be obtained through the integrating out of the Higgs bosons in every sector:
\begin{equation}\label{eqn:nnEffLag}
\begin{split}
\cL_\phi^{v \neq 0} &\supset C_1 a y_q\frac{v}{m_h^2}\phi q q^c,
\\
\cL_\phi^{v = 0} &\supset C_2 a \frac{g^2}{16\pi^2 m_H^2}\phi W_{\mu\nu}W^{\mu\nu},
\end{split}
\end{equation}
with $C_i$ representing numerical coefficients, $g$ the weak coupling constant, and $W^{\mu\nu}$ the SU$(2)$ field strength tensor.

Immediately from Eq. (\ref{eqn:nnEffLag}), we can see that the matrix element for decays into standard sectors is inversely proportional to that sectors higgs mass, $\cM_{m_H^2 < 0} \sim 1/m_{h_i}$. The Feynman diagram for this process is presented in Fig. \ref{fig:reheatonDecays}. The loop decay of $\phi \rightarrow \gamma\gamma$ is always sub-leading and can be neglected. It should be noted that as one goes to sectors with larger and larger vevs, the increasing mass of the fermions eventually leads to situations where the decay to two on-shell bottom or charm quarks is kinematically forbidden, $m_\phi < 2 m_q$. For sectors where this kinematic threshold is passed for charm quarks, contributions to cosmological observables can be safely ignored. All in all, we end up with a decay width that scales as $\Gamma_{m_H^2<0} \sim 1/m_h^2$. Since we can expect energy density to be proportional to the decay width, $\frac{\rho_i}{\rho_{SM}} \approx \frac{\Gamma_i}{\Gamma_{SM}}$, this indicates that energy density of standard sectors falls proportional to $1/i$.

For the exotic sectors, Eq. (\ref{eqn:nnEffLag}) indicates a matrix element scaling $\cM_{m_H^2>0} \sim 1/m_{H_i}^2$. As the Feynman diagram (Fig. \ref{fig:reheatonDecays}) shows, the reheaton decays to exotic sectors are loop suppressed, leading to a significantly lower energy density than the standard sectors. Both the decay width and energy density for these sectors scale as: $\Gamma_{m_H^2>0} \sim \rho_i \sim 1/m_H^4 \sim 1/i^2$.

As a final note, the reheating temperature, $T_{RH}$, has an upper bound on the order of the weak scale. If this bound is not observed, the SM Higgs mass would have major thermal corrections --- leading to the branching ratios into other sectors being problematically large. In accordance with this restriction, this project only examines reheating temperatures at or below $100$ GeV.

@@

As mentioned in the previous sections, for standard sectors with increasing index $i$ the vevs of said sectors increase $v_i\propto \sqrt{i}$. This leads to increasingly heavy particle spectra for higher sectors --- eventually leading to sectors that are essentially pure Yang-Mills that featuring strong first order phase transitions. This, of course, begs the question: at what index $i$ do said phase transitions begin?

@@

Using the methods outlined in the prior section we determine $\Lambda_{QCD}$ to have a relevant value of 
\begin{equation}\label{eqn:lambda2}
\Lambda^i_{(2)} = (m_s^i m_c^i m_b^i m_t^i)^{2/29}(\Lambda^i_{(6)})^{21/29}
\end{equation} 
at the energy scale we're interested in. $\Lambda^i_{(6)}$ is identical for all sectors and is taken to have a standard model value. Rewriting Eq. (\ref{eqn:lambda2}) in terms of standard model variables, 
\begin{equation}\label{eqn:lambda2adj}
\Lambda^i_{(2)} = (m_s m_c m_b m_t i^2)^{2/29}(\Lambda_{(6)})^{21/29}.
\end{equation}
Setting this equal to some multiple of the quark mass of sector $i$,
\begin{equation}
\kappa m^i_u = \kappa m_u \sqrt{i} = (m_s m_c m_b m_t i^2)^{2/29}(\Lambda_{(6)})^{21/29}.
\end{equation}
This can be solved for $i$:
\begin{equation}\label{eqn:critIndex}
i = \frac{(m_s m_c m_b m_t)^{4/25}(\Lambda_{(6)})^{42/25}}{(\kappa m_u)^{58/25}}.
\end{equation}

The critical index demonstrates that even in the most extreme cases, where first order phase transitions occur with the up quark an order of magnitude less than $\Lambda_{QCD}$, the critical index is still $\sim 1000$. Looking back to the description of reheating within $N$naturalness, we see the energy density of these sectors is at most $10^{-4}$ that of our sector. As a result, the gravitational waves generated by the standard large $i$ sectors in $N$naturalness will not generate detectable signatures.

However, if we move away from the standard $N$naturalness reheating mechanism and begin exploring mirror sectors with large vevs with $\cO( >10\%)$ of our energy density $\rho_{SM}$, we can have sectors with relatively high dark QCD scales that produce detectable gravitational waves. These sectors feature no baryons and, as such, produce runaway bubbles similar to the exotic sectors in standard $N$naturalness. From Eq. (\ref{eqn:lambda2}) we can determine the confinement scale of an arbitrary mirror sector --- the maximum possible given a $10^{16}$ limit \cite{Arkani-Hamed:2016rle} of sectors is $\sim 38 \,\gev$. After imposing $N_{eff}$ constraints (as in Sec. \ref{sec:constraints}), the energy density of said sector is limited to $\sim 0.25\rho_{SM}$. The signals of this sector and other test cases like it are explored in Sec. \ref{sec:signals}. It is important to note that these type of sectors mimic the general behaviour of any Yang-Mills like multi dark sector model.

@@

In every exotic sector the fermion masses are exceptionally light: even the most massive quark of each sector is orders of magnitude smaller than our confinement scale, $m_t \ll \Lambda_{QCD}$. Additionally, since $m_q \propto 1/(m^2_H)_i$ this separation only becomes more pronounced as we move to sectors further from our own. The scale of the exotic quarks makes it unnecessary to run down the scale of the strong coupling as done for the standard sectors --- the phase transition for all exotic sectors occurs at $\sim 90 \,\mathrm{MeV}$ and a strong first order phase transition occurs for all exotic sectors at this temperature.

The confinement of these sectors directly leads to the production of both baryons and mesons as we have the spontaneous breaking of SU$(6) \,\times$ SU$(6) \rightarrow\,$ SU$(6)$. This leads to the production of 35 pseudo-Goldstone bosons during the phase transition. These become relevant in the calculation of gravitational wave signatures through the calculation of the critical phase transition strength, $\alpha_{\infty}$, with $\alpha_{\infty}$ denotes the dividing line between the runaway regime  $(\alpha >\alpha_{\infty})$ and the non-runaway regime $(\alpha <\alpha_{\infty})$. Explicitly \cite{Breitbach:2018ddu, Caprini:2015zlo, Espinosa:2010hh}, 
\begin{equation}\label{eqn:critPTstrength}
\alpha_{\infty} = \frac{(T^{nuc}_h)^2}{\rho_R}\left[\sum_{bosons} n_i\frac{\Delta m^2_i}{24} + \sum_{fermions} n_i\frac{\Delta m_i^2}{48}\right],
\end{equation} 
for particles with $n_i$ degrees of freedom that obtain mass through the phase transition.

The masses obtained through the phase transition can be approximated through the use of a generalization of the Gell-Mann--Oakes--Renner relation \cite{Schwartz:2013pla},
\begin{equation}\label{eqn:gmor}
m^2_{\pi} = \frac{V^3}{F^2_{\pi}}(m_u + m_d),
\end{equation}
where $V \sim \Lambda_{QCD}$, $F_\pi$ is the pion decay constant. So, for mesons in exotic sector $i$:
\begin{equation}
m_{\pi}^i \sim \sqrt{\frac{m_1^i+m_2^i}{m_u + m_d}} m_{\pi}.
\end{equation}
Ultimately, calculating the critical phase transition strength for each sector using Eq. (\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Thus, all exotic sectors have strong first order phase transitions that lead to runaway bubble walls.

@@

The SM predicts that $N^{SM}_{eff} = 3.046$ \cite{Mangano:2005cc}. This is in good agreement with bounds from studies of the Cosmic Microwave Background (CMB) by Planck combined with bayon acoustic oscillations (BAO) measurements \cite{Aghanim:2018eyx}:
\begin{equation}\label{eqn:NeffBounds}
N_{eff} = 2.99^{+0.34}_{-0.33}.
\end{equation}
Various different assumptions about the history of the universe can be made and different data sets can be chosen to obtain slightly different results \cite{Breitbach:2018ddu} --- for the purposes of this exploratory work, wading through this landscape is unnecessary.

For fully decoupled sectors that never enters (or reenters) thermal equilibrium with our sector, we obtain additional contributions to $N^{SM}_{eff}$ \cite{Breitbach:2018ddu}
\begin{equation}\label{eqn:DeltaNeff1hs}
\Delta N_{eff} = \frac{4}{7}\left(\frac{11}{4}\right)^{4/3}g_h \xi_h^4.
\end{equation} 
Here, $g_h$ represents the effective number of relativistic degrees of freedom for the hidden sector when the hidden sector has a temperature $T_h = \xi_h T_\gamma$. Contours for Eq. (\ref{eqn:DeltaNeff1hs}) at various temperature ratios as well as CMB bounds are plotted in Fig. \ref{fig:Neff}. We take this approach and generalize it to include many additional sectors:
\begin{equation}\label{eqn:DeltaNeff}
\Delta N_{eff} = \sum_i \frac{4}{7}\left(\frac{11}{4}\right)^{4/3}g_{i} \xi^4_{i}.
\end{equation}

As Fig. \ref{fig:Neff} demonstrates, sectors with a temperature ratio above $\xi_h\sim 0.5$ with even a couple of relativist particles are immediately excluded by CMB constraints. Accordingly, the additional sectors present in this project must be heated to a much smaller fraction of our own sector's temperature.

@@

Within the framework of SM-like particle content (save the higgs mass parameter), our exotic sectors feature at most 106.75 effective degrees of freedom at high energies: 72 from quarks,  12 from charged leptons, 6 from neutrinos, 16 from gluons, 2 from photons, 6 from W bosons (no longitudinal modes due to a no symmetry breaking), 4 from the higgs sector (again, no symmetry breaking leads to 3 additional modes), for a total of 118 degress of freedom. Scaling the fermionic contributions by $7/8$ gives us 106.75 effective relativistic degrees of freedom.  After the higgses freeze out, 4 degrees of freedom are lost; when the QCD phase transition occurs 35 bosonic degrees of freedom are gained while 72 fermionic are lost, leaving 56.75 effective degrees of freedom. Since all additional sectors are colder than our own by the time photon decoupling occurs ($\sim 0.39 \,\ev$), the pions of all of these sectors will, in accordance with Eq. (\ref{eqn:gmor}), have a much higher mass than the sector temperature and thus will have frozen out long before --- leaving at most $11.25$ effective degrees of freedom per sector.

Coupling the number of effective degrees of freedom per sector with the energy density scaling of $\sim 1/m_H^4$ found in the zero vev sectors leads to tiny temperature ratios. Assuming a reheating temperature of $100$ GeV and a completely uniform distribution of sectors, the temperature of the first exotic sector is slightly more than $6\%$ of our sector at reheating. Applying Eq. (\ref{eqn:DeltaNeff}) to this particular situation gives us:
\begin{equation}
\Delta N_{eff} = \sum_i \frac{4}{7}\left(\frac{11}{4}\right)^{4/3}g_{i} \left(\frac{0.06}{i^2}\right)^4.
\end{equation}     
Summing over all sectors (sectors above $i \sim 10$ essentially don't contribute due to the small temperature ratios) gives us a contribution of $\mathcal{O} (10^{-4})$. Evolving the sectors thermal histories forward in time to the recombination era gives us a slightly larger value, but still of order $\mathcal{O}(10^{-4})$; well under current CMB bounds.

@@

Using these restrictions, we can examine the behaviour of standard sectors with a much larger vev than our own. In terms of the $N$naturalness framework, this consisted of looking at various different sectors above the critical index, as mentioned in Sec. \ref{sec:dQCD}. The signals for both the maximal case ($10^{16}$ limit) and several other high transition temperatures (involving sectors $i > 10^{10}$) scenarios are shown in Sec. \ref{sec:signals}.

>>>>

%\title{\LARGE{{\bf{Piled up gravitational waves:} \\
%\bf{Searching for new signals of $N$naturalness} }}}
\title{\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\
}}}
\author{{\bf {Paul Archer-Smith, Dylan Linthorne, and Daniel Stolarski}}}

@@

The recent experimental detection of gravitational waves~\cite{} gives humanity a new way to observe the universe. Future experiments~\cite{} will greatly expand the frequency range observable. Thus far, experiments have only observed recent events such as black hole mergers, but phase transitions in the early universe can leave an imprint as a \textit{stochastic} gravitational wave background~\cite{}. Thus searches for this background of gravitational waves can give direct information of the history of the universe before big bang nucleosynthesis.

Because gravity is universal, gravitational waves can allow us to probe hidden sectors that couple very weakly, or not at all, to the Standard Model as long they are reheated after inflation. This was first explored in~\cite{Schwaller:2015tja}, and there has been significant work on this idea since~\cite{}.

In this work, we explore the possibility of having multiple decoupled hidden sectors. Large numbers of hidden sectors can solve the hierarchy problem as in the Dvali Redi model~\cite{Dvali:2009ne} or in the more recently explored Nnaturalness~\cite{Arkani-Hamed:2016rle} framework, or in the orbifold Higgs models (maybe?\cite{}). They can also be motivated by dark matter considerations (dynamical dark matter? others?).

Motivated by solutions to the hierachy problem, we consider hidden sectors with the same particle content as the Standard Model that have all dimensionless couplings (defined at some high scale) equal to those of the Standard Model. The only parameter that varies across sectors is the dimension two Higgs mass squared parameter. This simple ansatz can lead to very rich phenomenology and interesting gravitational wave spectra, but we stress that it is only a starting point for exploring the space of theories with multiple hidden sectors.

\section{Particle Setup}
\label{sec:nn}

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=0.4\textwidth]{standardDecay.eps}
\includegraphics[width=0.4\textwidth]{exoticDecay.eps}
\end{minipage}
\hfill
\caption{Feynman diagrams for primary reheaton decays assuming $m_{\phi} \ll m_H$. The left decay is the primary decay for standard sectors ($\langle H\rangle\neq 0$), with $f$ representing all possible fermions. The right decay occurs in exotic sectors ($\langle H\rangle = 0$).
}
\label{fig:reheatonDecays}
\end{figure*}

@@

Keeping with the original convention, we write our mass parameter as
\begin{equation}\label{eqn:massParam}
\left(m_H^2\right)_i = -\frac{\Lambda_H^2}{N}(2i+r),
\end{equation}
where $i$ denotes the sector ($-\frac{N}{2}\leq i \leq \frac{N}{2}$). $i = 0$ is the sector that possess the smallest vev and is thus ``our" SM sector. \draftnote{Write an equation that shows how the vev scales with $i$.}Finally, the parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:signals}.

@@

A key issue within $N$naturalness is how to predominantly gift energy density to our own sector so as to not be immediately excluded by number of effective neutrinos ($N_{eff}$) bounds. This is done through the introduction of a post-inflationary field called the ``reheaton". After inflation, the reheaton field possess the majority of the energy density of the Universe. Although this field can generically be either by bosonic or fermionic, we reduce our scope to a scalar reheaton $\phi$. Our focus is primarily the production of gravitational waves from multiple sectors: \draftnote{What does the following mean?} the shifting to a fermion reheaton doesn't change the scaling of the energy density of the exotic sectors  and, as a result, is ignored in our analysis.

@@

Ultimately, after examining the gravitational wave case produced by standard $N$naturalness, we generalize our results to include the (rather unnatural) case where sectors are reheated ``randomly" as opposed to in relation to the higgs mass parameter. This allows us to explore more general models with multiple dark sectors at a huge range of scales. For these models, the reheating mechanism remains undefined and the results of this section are completely irrelevant.

The appropriate Lagrangian for a scalar reheaton $\phi$ is: 
\begin{equation}\label{eqn:nnLagrangian}
\cL_\phi \supset -a\phi\sum_i\vert H_i\vert^2 - \frac{1}{2}m^2_\phi \phi^2.
\end{equation}
Note that the cross-quartic couplings are absent, suppressed by a very small coupling. Effective Lagrangians for the two different types of sectors present in this theory can be obtained through the integrating out of the Higgs bosons in every sector:
\begin{equation}\label{eqn:nnEffLag}
\begin{split}
\cL_\phi^{v \neq 0} &\supset C_1 a y_q\frac{v}{m_h^2}\phi q q^c,
\\
\cL_\phi^{v = 0} &\supset C_2 a \frac{g^2}{16\pi^2 m_H^2}\phi W_{\mu\nu}W^{\mu\nu},
\end{split}
\end{equation}
\draftnote{Should the first equation go like $1/m_h$ (not squared)?}with $C_i$ representing numerical coefficients, $g$ the weak coupling constant, and $W^{\mu\nu}$ the SU$(2)$ field strength tensor.

Immediately from Eq. (\ref{eqn:nnEffLag}), we can see that the matrix element for decays into standard sectors is inversely proportional to that sectors higgs mass, $\cM_{m_H^2 < 0} \sim 1/m_{h_i}$. The Feynman diagram for this process is presented in Fig. \ref{fig:reheatonDecays}. The loop decay of $\phi \rightarrow \gamma\gamma$ is always sub-leading and can be neglected. It should be noted that as one goes to sectors with larger and larger vevs, the increasing mass of the fermions eventually leads to situations where the decay to two on-shell bottom or charm quarks is kinematically forbidden, $m_\phi < 2 m_q$. \draftnote{Put an equation sketching how $m_f$ scales with $m_h$} For sectors where this kinematic threshold is passed for charm quarks, contributions to cosmological observables can be safely ignored. All in all, we end up with a decay width that scales as $\Gamma_{m_H^2<0} \sim 1/m_h^2$. Since we can expect energy density to be proportional to the decay width, $\frac{\rho_i}{\rho_{SM}} \approx \frac{\Gamma_i}{\Gamma_{SM}}$, this indicates that energy density of standard sectors falls proportional to $1/i$.

For the exotic sectors, Eq. (\ref{eqn:nnEffLag}) indicates a matrix element scaling $\cM_{m_H^2>0} \sim 1/m_{H_i}^2$. As the Feynman diagram (Fig.~\ref{fig:reheatonDecays}) shows, the reheaton decays to exotic sectors are loop suppressed, leading to a significantly lower energy density than the standard sectors. Both the decay width and energy density for these sectors scale as: $\Gamma_{m_H^2>0} \sim \rho_i \sim 1/m_H^4 \sim 1/i^2$.

As a final note, the reheating temperature, $T_{RH}$, \draftnote{What is this parameter exactly? The reheat temperature of the SM? Also, is this effect discussed in the original NNaturalness paper?} has an upper bound on the order of the weak scale. If this bound is not observed, the SM Higgs mass would have major thermal corrections --- leading to the branching ratios into other sectors being problematically large. In accordance with this restriction, this project only examines reheating temperatures at or below $100$ GeV.

@@

As mentioned in the previous sections,\draftnote{refer to a particular equation} for standard sectors with increasing index $i$ the vevs of said sectors increase $v_i\propto \sqrt{i}$. This leads to increasingly heavy particle spectra for higher sectors --- eventually leading to sectors that are essentially pure Yang-Mills that featuring strong first order phase transitions. This, of course, begs the question: at what index $i$ do said phase transitions begin?

@@

Using the methods outlined in the prior section we determine $\Lambda_{QCD}$ to have a relevant value of 
\begin{equation}\label{eqn:lambda2}
\Lambda^i_{(2)} = (m_s^i m_c^i m_b^i m_t^i)^{2/29}(\Lambda^i_{(6)})^{21/29}
\end{equation} 
at the energy scale we're interested in. $\Lambda^i_{(6)}$ is identical for all sectors and is taken to have a standard model value \draftnote{what is that value?}. Rewriting Eq. (\ref{eqn:lambda2}) in terms of standard model variables, 
\begin{equation}\label{eqn:lambda2adj}
\Lambda^i_{(2)} = (m_s m_c m_b m_t i^2)^{2/29}(\Lambda_{(6)})^{21/29}.
\end{equation}
Setting this equal to some multiple of the quark mass of sector $i$,
\begin{equation}
\kappa m^i_u = \kappa m_u \sqrt{i} = (m_s m_c m_b m_t i^2)^{2/29}(\Lambda_{(6)})^{21/29}.
\end{equation}
This can be solved for $i$:
\begin{equation}\label{eqn:critIndex}
i = \frac{(m_s m_c m_b m_t)^{4/25}(\Lambda_{(6)})^{42/25}}{(\kappa m_u)^{58/25}}.
\end{equation}

The critical index demonstrates that even in the most extreme cases, where first order phase transitions occur with the up quark an order of magnitude less than $\Lambda_{QCD}$, the critical index is still $\sim 1000$. Looking back to the description of reheating \draftnote{refer to a particular equation} within $N$naturalness, we see the energy density of these sectors is at most $10^{-4}$ \draftnote{why not $10^{-3}$?} that of our sector. As a result, \draftnote{As we will show below?} the gravitational waves generated by the standard large $i$ sectors in $N$naturalness will not generate detectable signatures.

However, if we move away from the standard $N$naturalness reheating mechanism and begin exploring mirror sectors with large vevs with $\cO( >10\%)$ of our energy density $\rho_{SM}$, we can have sectors with relatively high dark QCD scales that produce detectable gravitational waves. These sectors feature no baryons and, as such, produce runaway bubbles similar to the exotic sectors in standard $N$naturalness. From Eq. (\ref{eqn:lambda2}) we can determine the confinement scale of an arbitrary mirror sector --- the maximum possible given a $10^{16}$ limit \cite{Arkani-Hamed:2016rle} \draftnote{Is this actually an upper value of $m_H^2$ at the gut scale?} of sectors is $\sim 38 \,\gev$. After imposing $N_{eff}$ constraints (as in Sec. \ref{sec:constraints}), the energy density of said sector is limited to $\sim 0.25\rho_{SM}$. The signals of this sector and other test cases like it are explored in Sec. \ref{sec:signals}. It is important to note that these type of sectors mimic the general behaviour of any Yang-Mills like multi dark sector model. \draftnote{What does this sentence mean? Are there references that should appear here?}

@@

In every exotic sector the fermion masses are exceptionally light: their masses are generated by dimension six operators with the Higgs integrated out as show in \draftnote{cite an equation from before}, and are therefore all below the confinement scale. The exotic sectors all have identical one-loop running of the QCD gauge coupling, and thus all have approximately the same confinement scale given by $\Lambda_{\rm ex} \sim 90 \,\mathrm{MeV}$. These sectors all have six light fermions, so a strong first order phase transition occurs for all exotic sectors at this temperature.

The confinement of these sectors directly leads to the production of both baryons and mesons as we have the spontaneous breaking of SU$(6) \,\times$ SU$(6) \rightarrow\,$ SU$(6)$ and thus 35 pseudo-Goldstone bosons. These become relevant in the calculation of gravitational wave signatures through the calculation of the critical phase transition strength, $\alpha_{\infty}$, with $\alpha_{\infty}$ denotes the dividing line between the runaway regime  $(\alpha >\alpha_{\infty})$ and the non-runaway regime $(\alpha <\alpha_{\infty})$. Explicitly \cite{Breitbach:2018ddu, Caprini:2015zlo, Espinosa:2010hh}, 
\begin{equation}\label{eqn:critPTstrength}
\alpha_{\infty} = \frac{(T^{nuc}_h)^2}{\rho_R}\left[\sum_{bosons} n_i\frac{\Delta m^2_i}{24} + \sum_{fermions} n_i\frac{\Delta m_i^2}{48}\right],
\end{equation} 
for particles with $n_i$ degrees of freedom that obtain mass through the phase transition.

The masses obtained through the phase transition can be approximated through the use of a generalization of the Gell-Mann--Oakes--Renner relation \draftnote{Also cite Gell-Mann et al} \cite{Schwartz:2013pla},
\begin{equation}\label{eqn:gmor}
m^2_{\pi} = \frac{V^3}{F^2_{\pi}}(m_u + m_d),
\end{equation}
where $V \sim \Lambda_{QCD}$, $F_\pi$ is the pion decay constant. So, for mesons in exotic sector $i$:
\begin{equation}
m_{\pi}^i \sim \sqrt{\frac{m_1^i+m_2^i}{m_u + m_d}} m_{\pi}.
\end{equation}
\draftnote{What are $m_a^i$? Is this the right formula,isn't $V$ and $f_\pi$ different from the SM to the $i$th exotic sector? Also, how do you compute $\Delta m$ when you have a different set of particles inside (pions) as outside (quarks) the bubble?  } Ultimately, calculating the critical phase transition strength for each sector using Eq. (\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Thus, all exotic sectors have strong first order phase transitions that lead to runaway bubble walls.

@@

\draftnote{This is not the physics described in the previous paragraph} The SM predicts that $N^{SM}_{eff} = 3.046$ \cite{Mangano:2005cc}. This is in good agreement with bounds from studies of the Cosmic Microwave Background (CMB) by Planck combined with bayon acoustic oscillations (BAO) measurements \cite{Aghanim:2018eyx}:
\begin{equation}\label{eqn:NeffBounds}
N_{eff} = 2.99^{+0.34}_{-0.33}.
\end{equation}
\draftnote{Are these $1\sigma$? Maybe show the 95\% interval instead?}
Various different assumptions about the history of the universe can be made and different data sets can be chosen to obtain slightly different results \cite{Breitbach:2018ddu} --- for the purposes of this exploratory work, wading through this landscape is unnecessary.

For fully decoupled sectors that never enter (or reenter) thermal equilibrium with our sector, we obtain additional contributions to $N^{SM}_{eff}$ \cite{Breitbach:2018ddu}
\begin{equation}\label{eqn:DeltaNeff1hs}
\Delta N_{eff} = \frac{4}{7}\left(\frac{11}{4}\right)^{4/3}g_h \xi_h^4.
\end{equation} 
Here, $g_h$ represents the effective number of relativistic degrees of freedom for the hidden sector when the hidden sector has a temperature $T_h = \xi_h T_\gamma$. Contours for Eq. (\ref{eqn:DeltaNeff1hs}) at various temperature ratios as well as CMB bounds are plotted in Fig.~\ref{fig:Neff}. We take this approach and generalize it to include many additional sectors:
\begin{equation}\label{eqn:DeltaNeff}
\Delta N_{eff} = \sum_i \frac{4}{7}\left(\frac{11}{4}\right)^{4/3}g_{i} \xi^4_{i}.
\end{equation}

As Fig.~\ref{fig:Neff} \draftnote{Why is it figure V?} demonstrates, sectors with a temperature ratio above $\xi_h\sim 0.5$ with even a couple of relativistic particles are immediately excluded by CMB constraints. Accordingly, the additional sectors present in this project must be heated to a much smaller fraction of our own sector's temperature. \draftnote{$\xi=0.5$ and $g=2$ is not excluded.}

@@

Within the framework of SM-like particle content (save the higgs mass parameter), our exotic sectors feature at most 106.75 effective degrees of freedom at high energies: 72 from quarks,  12 from charged leptons, 6 from neutrinos, 16 from gluons, 2 from photons, 6 from W bosons (no longitudinal modes due to a no symmetry breaking), 4 from the higgs sector (again, no symmetry breaking leads to 3 additional modes), for a total of 118 degress of freedom. Scaling the fermionic contributions by $7/8$ gives us 106.75 effective relativistic degrees of freedom.  After the higgses freeze out, 4 degrees of freedom are lost; when the QCD phase transition occurs 35 bosonic degrees of freedom are gained while 72 fermionic are lost, leaving 56.75 effective degrees of freedom. Since all additional sectors are colder than our own \draftnote{Aren't they always colder than our own?} by the time photon decoupling occurs ($T_\gamma \sim 0.39 \,\ev$), the pions of all of these sectors will, in accordance with Eq. (\ref{eqn:gmor}), have a much higher mass than the sector temperature and thus will have frozen out long before --- leaving at most $11.25$ effective degrees of freedom per sector \draftnote{From photons and neutrinos? What about $\Delta N_{eff}$ constraints from BBN? This would be a much higher (SM) temperature.} .

Coupling the number of effective degrees of freedom per sector with the energy density scaling of $\sim 1/m_H^4$ found in the zero vev sectors leads to tiny temperature ratios. Assuming a reheating temperature of $100$ GeV and a completely uniform distribution of sectors, the temperature of the first exotic sector is slightly more than $6\%$ of our sector at reheating. Applying Eq.~(\ref{eqn:DeltaNeff}) to this particular situation gives us:
\begin{equation}
\Delta N_{eff} = \sum_i \frac{4}{7}\left(\frac{11}{4}\right)^{4/3}g_{i} \left(\frac{0.06}{i^2}\right)^4 \sim 10^{-4}.
\end{equation}     
\draftnote{Where does the $0.06$ come from?} This is dominated by $i=1$, the sector with the lowest Higgs mass (and thus the most energy density) gives us a contribution of $\mathcal{O} (10^{-4})$. Evolving the sectors thermal histories forward in time to the recombination era gives us a slightly larger value, but still of order $\mathcal{O}(10^{-4})$; well under current CMB bounds.

@@

Using these restrictions, we can examine the behaviour of standard sectors with a much larger vev than our own. In terms of the $N$naturalness framework, this consisted of looking at various different sectors above the critical index, as mentioned in Sec. \ref{sec:dQCD}. The signals for both the maximal case ($10^{16}$ limit) and several other high transition temperatures (involving sectors $i > 10^{10}$) scenarios are shown in Sec. \ref{sec:signals}. \draftnote{This assumes that only photons and neutrinos are relativistic?}

\section{Gravitational Wave Signals}

\label{sec:gw}
Sectors other than the SM can be reheated through multifield inflationary theories, Reheaton models, etc. In these models, the SM will receive the majority of the energy density while the remaining energy density will be distributed, in a model dependent way, to the various hidden sectors. If there does not exist a portal between the sectors, they will remain decoupled and evolve independently. When looking at multiple thermally decoupled hidden sectors it is convienent to define the temperature ratio [cite cold,dark,noisy]
\begin{equation}
\xi_i \equiv  \frac{T_{i}}{T_{\gamma}}
\end{equation}
Where $T_{i}$ is the temperature of the $i^{th}$ hidden sector, and $T_{\gamma}$ the temperature of the standard model (assuming radiation domination).
